How to be a good member of a scientific software community [Article v\versionnumber]

### 2021-08-28 paper.tex
incorporated some of Tod's ideas

Software is ubiquitous in modern science --- almost any project, in almost any
discipline, requires some code to work. However, many (or even most) scientists
are not programmers, and must rely on programs written and maintained by others.
As a result, a crucial but often neglected part of a scientist's training is
learning how to use new tools, and how to exist as part of a community of users.
This article will discuss key behaviors that can make the experience quicker,
more efficient, and more pleasant for the user and developer alike.

@@

Most practicing scientists (even the ones who write code as part of their
research) spend more time as consumers as opposed to producers of software.
Moreover, we are constantly learning new skills and solving new problems, which
often means learning to use new programs or new aspects of large packages. This,
combined with the complex nature of scientific software (and the often low
quality of the associated documentation) means we have to ask for help.
Sometimes it's because we don't know how to accomplish a specific task. Others,
it's because there's a feature we need that isn't implemented. Inevitably, there
are bugs.

In all of these cases, it is necessary to interact with the people who develop
and support the code. How you go about it has an enormous impact on your
likelihood of success and how you are viewed. Asking complete strangers for
help is often intimidating, especially the first time you do it. However, the
best way to do so is rarely taught explicitly --- at best, it's something one
picks up by example, from fellow lab members or the PI.  {\emph The goal of this
paper is to reveal the hidden curriculum -- what's expected of a software
consumer, how to ask for help, how to contribute productively to a software
community (regardless of whether you can write code).}

@@

The second step is to check the manual. This might seem trivially obvious, but
you'd be astonished how many emails developers get that are directly answered in
the manual. If the manual is short, you can read the whole thing. If it's
longer, search for relevant-seeming keywords, scan the sections that seem to
pertain to your error message. If there's an FAQ (frequently asked questions)
list, check that first; asking an FAQ is likely to lead to grumbling.

@@

@@

    \item Describe the platform you're running on (e.g. the operating system for
    a conventional computer, hardware if it's a tablet, etc.). If the problem is
    related to building or installing the software, describe the build system
    (e.g. compiler version) or package manager (e.g. Conda\cite{ANACONDA}).

@@

Once you've become reasonably competent with a piece of software, it's worth
looking for ways to contribute back. This can take many forms, not all of which
involve writing code (though of course contributing bug fixes or new features is
extremely valuable). You can work on the documentation --- if you struggled to
understand something but eventually figured it out, consider contributing a new
paragraph of text; many packages hosted on GitHub have the documentation in the
same repository, so the same pull request mechanism can be used, but even if the
docs are separate you could write a new paragraph or two and send them to the
developer. If you developed a toy example as you were figuring out how to use
the software, consider sharing it. If one of the provided examples had to be
tweaked to work, share the tweaks. If the software has an online forum (e.g.
mailing list, GitHub discussions or issues, etc), you might be able to
contribute by suggestion solutions to problems from less experienced users.
However, unlike the other recommentations, this one isn't mandatory --- it's a
``nice to have'' rather than a ``must have''.

@@

>>>>

Software is ubiquitous in modern science --- almost any project, in almost any
discipline, requires some code to work. However, many (or even most) scientists
are not programmers, and must rely on programs written and maintained by others.
A crucial but often neglected part of a scientist's training is
learning how to use new tools, and how to exist as part of a community of users.
This article will discuss key behaviors that can make the experience quicker,
more efficient, and more pleasant for the user and developer alike.

@@

Most practicing scientists (even the ones who write code as part of their
research) spend more time as consumers as opposed to producers of software.
Moreover, we are constantly learning new skills and solving new problems, which
often means learning to use new programs or new aspects of large packages. This,
combined with the complex nature of scientific software (and often the missing
or inaccurate documentation) means we have to ask for help.  Sometimes it's
because we don't know how to accomplish a specific task. Others, it's because
there's a feature we need that isn't implemented. Inevitably, there are bugs.

In all of these cases, it is necessary to interact with the people who develop
and support the code. How you go about it has an enormous impact on your
likelihood of success and how you are viewed. Asking complete strangers for help
is often intimidating, especially the first time you do it. However, the best
way to do so is rarely taught explicitly --- at best, it's something one picks
up by example, from fellow lab members or your PI.  {\emph The goal of this
paper is to reveal the hidden curriculum -- what's expected of a software
consumer, how to ask for help, how to contribute productively to a software
community (regardless of whether you can write code).}

@@

The second step is to check the manual. This might seem trivially obvious, but
you'd be astonished how many emails developers get that are directly answered in
the manual. If the manual is short, you can read the whole thing. If it's
longer, search for relevant-seeming keywords, scan the sections that seem to
pertain to your error message. If there's an FAQ (frequently asked questions)
list, check that first; asking an FAQ, especially in a public forum, is likely
to lead to grumbling.

@@

One valuable strategy is to construct a minimal set of circumstances that causes
the problem. Even if you initially encountered the problem while analyzing a 2
TB data set, that's not going to be very helpful for diagnosing the problem (and
you won't easily be able to share the data set with the developers so they can
diagnose it). Can you reproduce the problem using only a small portion of the
data? Is the problem present with some kinds of data (e.g. some frames of the
trajectory) but not others, and if so what is different about those frames? Can
you make the problem appear and disappear with small changes in the settings?
Testing these things out may lead you to a solution, but even if they don't
they'll help you describe the problem more precisely to the developers.

@@

    \item Describe the platform you're running on (e.g. the operating system for
    a conventional computer, hardware if it's a tablet, etc.). If the problem is
    related to building or installing the software, describe the build system
    (e.g. compiler version) or package manager (e.g. Conda\cite{CONDA}, Homebrew \cite{HOMEBREW}).

@@

Once you've become reasonably competent with a piece of software, it's worth
looking for ways to contribute back. This can take many forms, not all of which
involve writing code (though of course contributing bug fixes or new features is
extremely valuable). You can work on the documentation --- if you struggled to
understand something but eventually figured it out, consider contributing a new
paragraph of text. Many packages hosted on GitHub have the documentation in the
same repository, so the same pull request mechanism can be used, but even if the
docs are separate you could write a new paragraph or two and send them to the
developer. If you developed a toy example as you were figuring out how to use
the software, consider sharing it. If one of the provided examples had to be
tweaked to work, share the tweaks. If the software has an online forum (e.g.
mailing list, GitHub discussions or issues, etc), you might be able to
contribute by answering questions from less experienced users.

Unlike the other recommentations, this one isn't mandatory --- it's a ``nice to
have'' rather than a ``must have''. However, there are a number of advantages to
doing it. First and foremost, you'll be making a package that's valuable to you
stronger, which will help you with your own scientific projects. However,
participating productively in a software community is also a good way to build
your scientific communication skills. Finally, helping others is an excellent
way for a relatively junior scientist to build their reputation in the
community, which in turn can lead to future collaborations or even job
opportunities. At the very least, it's always valuable when a lot of people in
your field have positive associations with your name.

@@

If the justice perspective doesn't persuade you, consider the practical
consequences. Bad behavior on the part of developers of your project (or your
community in general) will drive users and developers away from you. If
interacting with your community is unpleasant, folks will either use alternative
codes or use your code but not cite you. The next person with an idea for an
innovative new feature will add it to a different package, or won't do it at
all (hurting the whole community). Ask yourself: can your project really afford to drive away talented hard-working people eager to help? Is it worth is, just so a few people don't have to engage with basic courtesy?
